---
title: "Downstream analysis followig the image analysis to quanttify Dvl2_mEos codensates"
subtitle: "Markdown to reproduce Figure 7B"
author:  
- name: "Christian Scheeder"
- name: "Antionia Schubert"
  affiliation:
  - "Division Signalling and Functional Genomics, German Cancer Research Center (DKFZ) Heidelberg"  
date: \today
output:
    BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
editor_options: 
    chunk_output_type: console
---



# Preamble
The image analysis was carried out as follows:
In a first step cells were identified by segmentation. Nuclei were identified 
based on a Hoechst staining (DAPI channel) and cell bodies 
based on a DyLight Phalloidin staining 
(Cy5 channel). 
After segmentation condensates inside cells were identified based on intensity-
based thresholding (FITC channel) and counted. Condensate counts and 
cell counts (number of segmented cells) were saved for each image.
For more details see Schubert et al., PNAS, 2021, 
(https://doi.org/10.1073/pnas.2122476119)


# Load dependencies
```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(cowplot)
library(ggpubr)
theme_set(theme_cowplot(font_size = 12))
```


# Load the data after the image analysis
Note: Per well, four field were imaged and analyzed.
      For subsequent analysis the data is averaged per well
```{r}
files <- list.files(here("raw_data/figure7"))

raw_data <- lapply(files, function(x)
  read_rds(file.path(here("raw_data/figure7"),x))) %>% 
  bind_rows()
              
# average across the fields 
raw_data <- raw_data %>% separate(well_id,c("well_id","field_id"),"_") %>% 
  group_by(experiment,plate_id,well_id) %>% 
        summarise_if(is.numeric,median) %>% 
          ungroup() 

# add annotation
anno <- read_csv(here("annotation/figure7","figure7_annotation.csv")) %>% 
          select(well_ID,Cell_line,Treatment) %>% 
          `colnames<-`(c("well_id", "cell_line", "treatment"))


raw_data <- raw_data %>% 
  left_join(anno,by=c("well_id"))

```


# Normalize condensate counts 
Condensate counts are normalized to cell count 
(i.e. the number of segmented cells) and average per plate
```{r}
norm_data <- raw_data %>% 
  select(well_id,experiment,plate_id,
         cell_count,punct_count_adapt,punct_count_global,
         cell_line,treatment) %>% 
   mutate(norm_punct_count = punct_count_adapt/cell_count) %>%
    group_by(experiment,plate_id,cell_line,treatment) %>%
    summarise_if(is.numeric,median) %>% 
    ungroup() %>% 
      select(-cell_count,-punct_count_global,-punct_count_adapt)
```


# Plot for Figure 7 B 
```{r}
set.seed(1)
boxplot_medium <- norm_data %>% 
        filter(treatment == "medium") %>%
        mutate(cell_line = factor(cell_line,
                                  levels = c("HEK293T_Dvl2_mEos",
                                             "HEK293T_Dvl2_mEos_EVI_KO",
                                             "HEK293T_Dvl2_mEos_MCC_KO",
                                             "HEK293T_Dvl2_mEos_APC_trunc",
                                             "HEK293T_Dvl2_mEos_APC_KO_IV",
                                             "HEK293T_Dvl2_mEos_APC_KO_V",
                                             "HEK293T_Dvl2_mEos_Axin_KO_IV",
                                             "HEK293T_Dvl2_mEos_Axin_KO_V",
                                             "HEK293T_Dvl2_mEos_Axin_KO_VI",
                                             "HEK293T_Dvl2_mEos_Axin_KO_VII",
                                             "HEK293T_wt"),
                                  labels = c("HEK293T_Dvl2_mEos",
                                             "HEK293T_Dvl2_mEos_EVI_KO",
                                             "HEK293T_Dvl2_mEos_MCC_KO",
                                             "HEK293T_Dvl2_mEos_APC_trunc",
                                             "HEK293T_Dvl2_mEos_APC_KO_IV",
                                             "HEK293T_Dvl2_mEos_APC_KO_V",
                                             "HEK293T_Dvl2_mEos_Axin_KO_IV",
                                             "HEK293T_Dvl2_mEos_Axin_KO_V",
                                             "HEK293T_Dvl2_mEos_Axin_KO_VI",
                                             "HEK293T_Dvl2_mEos_Axin_KO_VII",
                                             "HEK293T_wt"))) %>% 
        ggplot(aes(x = cell_line,y=norm_punct_count)) + 
        ggtitle("Condition: Medium") + 
                geom_boxplot(outlier.shape = NA) + 
                geom_jitter() +
    ylab(
"normalized condensate counts") +
    xlab("condition") + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  theme(plot.title = element_text(hjust = 0.5))


ggsave(here("./graphics/Figure7b","figure7b.pdf"),
            plot=boxplot_medium,
            width = 14,height = 10)

```


# Session info
```{r}
sessionInfo()
```