---
title: "Downstream analysis followig the image analysis to quanttify Dvl2_mEos codensates"
subtitle: "Markdown to reproduce Figure 7F"
author:  
- name: "Christian Scheeder"
- name: "Antionia Schubert"
  affiliation:
  - "Division Signalling and Functional Genomics, German Cancer Research Center (DKFZ) Heidelberg"  
date: \today
output:
    BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
editor_options: 
    chunk_output_type: console
---


# Preamble
The image analysis was carried out as follows:
In a first step cells were identified by segmentation. Nuclei were identified 
based on a Hoechst staining (DAPI channel) and cell bodies 
based on a DyLight Phalloidin staining 
(Cy5 channel). 
After segmentation condensates inside cells were identified based on intensity-
based thresholding (FITC channel) and counted. Condensate counts and 
cell counts (number of segmented cells) were saved for each image.
For more details see Schubert et al., PNAS, 2021, 
(https://doi.org/10.1073/pnas.2122476119)


# Load dependencies
```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(cowplot)
library(ggpubr)
library(reshape2)
library(pheatmap)
theme_set(theme_cowplot(font_size = 12))
```




# Load the data after the image analysis
```{r}
files <- list.files(here("raw_data/figure7"))

raw_data <- lapply(files, function(x)
  read_rds(file.path(here("raw_data/figure7"),x))) %>% 
  bind_rows()
              
# average across the fields 
raw_data <- raw_data %>% separate(well_id,c("well_id","field_id"),"_") %>% 
  group_by(experiment,plate_id,well_id) %>% 
        summarise_if(is.numeric,median) %>% 
          ungroup() 

# add annotation 
anno <- read_csv(here("annotation/figure7","figure7_annotation.csv")) %>% 
          select(well_ID,Cell_line,Treatment) %>% 
          `colnames<-`(c("well_id", "cell_line", "treatment"))


raw_data <- raw_data %>% 
  left_join(anno,by=c("well_id"))

```


# Normalize condensate counts 
Condensate counts are normalized to cell count 
(i.e. the number of segmented cells) and average per plate
```{r}
norm_data <- raw_data %>% 
  select(well_id,experiment,plate_id,cell_count,punct_count_adapt,punct_count_global,
         well_id,cell_line,treatment) %>% 
   mutate(norm_punct_count = punct_count_adapt/cell_count) %>%
    group_by(experiment,plate_id,cell_line,treatment) %>%
    summarise_if(is.numeric,median) %>% 
    ungroup() %>% 
      select(-cell_count,-punct_count_global,-punct_count_adapt)
```



# Morphology analysis
Note: In addition to the condensate and cell counts, morphological features 
   describing e.g. cell size and shape extracted. 
   Features were extracted of single cells and then averaged per image.
   Those image-level averages were then further averaged per well and 
   plate as described above. The image analysis and feature extraction 
   procedure is described in detail in Schubert et al. 2021
   (https://doi.org/10.1073/pnas.2122476119)

```{r}
anno_cols <- c("well_id","experiment","plate_id","cell_line","treatment")

               
# bring data in long format 
df_long <- raw_data %>% melt(
  id.vars=c(anno_cols),
  variable.name = "feature",
  value.name = "measurement") %>% 
  mutate(feature = as.character(feature))


# drop features from the CFITC channel 
#   (i.e. the channel which was used to detect mEos)
df_long <- df_long %>% filter(!grepl("tub",feature))


# transform (generalized logarithm) and scale (robust z-score) the data 
glog <-
function(x, a=1, InverseQ=FALSE) {
if (InverseQ) {
    out<-0.25*exp(-x)*(4*exp(2*x)-(a*a))
}
else
    out<-log((x + sqrt(x^2 + a^2))/2)
out
}


df_long <- df_long %>% group_by(feature) %>% 
  mutate(measurement = glog(measurement)) %>% 
  mutate(measurement = 
      (measurement - median(measurement,na.rm=T))/mad(measurement,na.rm=T)) %>% 
  ungroup()




# we use a defined set of features for morphology analysis:
# mean_cell.0.m.eccentricity: average eccentricity of cells 
# mean_cell.0.s.area: average area of cells 
# mean_nuc.0.s.area: average area of nuceli 

kept_features <- c("mean_cell.0.m.eccentricity",
                   "mean_nuc.0.s.area",
                   "mean_cell.0.s.area",
                   "cell_count")


feature_matrix <- df_long %>%
  filter(treatment != "medium") %>% 
  filter(cell_line != "HEK293T_wt") %>% 
  filter(feature %in% kept_features) %>% 
  group_by(cell_line,treatment,feature) %>% 
  summarise(measurement = median(measurement,na.rm=T)) %>% 
  ungroup() %>% 
  mutate(cell_line = str_replace(cell_line,"HEK293T_","")) %>% 
  unite(id,cell_line,treatment) %>% 
  spread(feature,measurement) %>%
  column_to_rownames("id")


features_breakPoints <- c(min(feature_matrix),
                          seq(-2,2,by=0.01),
                          max(feature_matrix))

features_cols <- colorRampPalette(
  c("#a1d76a","#f7f7f7","#e9a3c9"))(length(features_breakPoints))

feature_matrix_heatmap <- pheatmap(feature_matrix,
                                   color = features_cols,
                                             )

pdf(
  here(
    "./graphics/figure7f",
    "figure7f.pdf"),
  width=10, height=8)
feature_matrix_heatmap
dev.off()

```


# Session info
```{r}
sessionInfo()
```
