---
title: "Down_stream_analysis"
author: "Christian Scheeder, Florian Heigwer, Antonia Schubert"
date: \today
output: html_document
editor_options: 
    chunk_output_type: console
---

# Load dependencies

```{r setup, include=FALSE,echo=FALSE,message=F,warning=F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggpubr)
library(RColorBrewer)
library(reshape2)
library(FitAR)
library(ggrepel)
library(caret)
library(pheatmap)
library(heatmaply)
library(DT)
library(cowplot)
library(ggrastr)
theme_set(theme_cowplot())
```

# Preamble

In this experiment cell lines possessing a endogenously tagged Dsh allele were subjected to a high-throughput compound screening experiment. To this end, different cell lines were seeded, and pre-incubated at 25 °C for 24 h. Afterwards cells were treated with 0.5 µM or 5 µM of chemical compound (concentration "l" and "h" respectively). After treatment cells were incubated for further 48 h and afterwards fixed and stained for DNA, actin and alpha-tubulin. Cells were then imaged using the InCell 2200 Analyzer at 20 x magnification, 3 channels per field of view, 4 fields per well. Cells were then segmented from each image and analyzed for morphology and marker staining intensity.

During analysis it's ought to be found out if there are chemical compounds or repurposed drugs that decrease or increase the abundance of Dsh condensates throughout the cytoplasm. So the data is read int, filtered, glog transformed, quantile normalized, centered and scaled per screen, plate and feature and then scored for reproducibility and foldchange in condensate abundance.

cl080 = HEK293 wild-type 
cl081 = HEK293 (Dsh-eosFP)
cl082 = other

# Load the data 
# The data set was pulled from the data base and average per well 

(mean over four fields per well, see get_data.Rmd for details).

```{r,echo=F,warning=F,message=F}
raw_data <- readRDS(here("data","HC1098_S01_rawData_sum.rds")) 
```

# Add the cell line annotation with redaily interpretable names

```{r,echo=F,warning=F,message=F}
cell_line_anno <- 
  tibble(cell_line_id=c("cl080","cl081","cl082"),
             cell_line = c("wildtype",
                           "Dsh_eos",
                           "FZD_ko"))

raw_data <- raw_data %>% 
  left_join(cell_line_anno,by=c("cell_line_id")) %>% 
  filter(cell_line != "FZD_ko") 

```

# Add library annotations 

```{r,echo=F,warning=F,message=F}

# load the annotation files 
anno_inf <- read_delim(here("annotation/old","informerv2_anno_final.txt"), delim = ",") %>% 
          mutate(library_plate_id = "1",library_id = "INF1") %>%
          rename(target_cc=target)%>%
          mutate(target_selleck=target_cc)

anno_kin <- read_delim(here("annotation","HC1094_selleckKIN1_anno.csv"), delim = ",") %>% 
          mutate(library_plate_id = as.character(plate_id),library_id = "KIN1")%>%
          extract(library_plate_id,"library_plate_id",".+_p(\\d)") 

anno_FDA <- read_delim(here("annotation","HC1094_selleckFDA1_anno.csv"), delim = ",") %>% 
          mutate(library_plate_id = as.character(plate_id),library_id = "FDA1")%>%
          extract(library_plate_id,"library_plate_id",".+_p(\\d)") 

anno <- bind_rows(anno_inf,anno_kin,anno_FDA)

anno %<>%
  separate(target_cc,into = c("temp1","temp2","temp3"),extra = "drop",sep = ";") %>%
  mutate(temp2=if_else(is.na(temp2),"",temp2),
         temp3=if_else(is.na(temp3),"",temp3)) %>%
  unite(temp1,temp2,temp3,col = "target_cc",sep = ";") %>%
  mutate(target_cc=gsub(";+$","",target_cc)) %>%
  mutate(target_cc=if_else(grepl("NA",target_cc),if_else(!grepl("DMSO",target_selleck),target_selleck,"NA"),target_cc)) %>%
   mutate(value=if_else(grepl("DMSO",product_name),"DMSO_ctrl",value)) %>%
  mutate(target_cc=if_else(grepl("DMSO",product_name),"DMSO_ctrl",target_cc)) %>%
  mutate(well_id = gsub("^(\\w)(\\d)$","\\10\\2",well_id))

# left join annotation and data 
raw_data <- raw_data %>% 
  mutate(library_plate_id = as.character(library_plate_id)) %>% 
  left_join(anno,by=c("well_id","library_plate_id","library_id")) %>% 
  select(-fussel,-dark,-dead_cells)

```

# Manually annotate the positive controls and change the annotation for the FZD ko

```{r,echo=F,warning=F,message=F}

raw_data <- raw_data %>% 
  mutate(value="sample") %>%
  mutate(value = if_else(grepl("DMSO",target_selleck), "DMSO_ctrl", value)) %>% 
  mutate(type = if_else(grepl("DMSO",target_selleck), "DMSO_ctrl", type)) %>%
  mutate(product_name = if_else(grepl("DMSO",target_selleck), "DMSO_ctrl", product_name)) %>% 
  mutate(catalog_number = if_else(grepl("DMSO",target_selleck), "DMSO_ctrl", catalog_number)) %>% 
  mutate(target_cc = if_else(grepl("DMSO",target_selleck), "DMSO_ctrl", target_cc))

pos_ctrls <- c("B02","G12","M21","L07","E17")

raw_data <- raw_data %>% 
  mutate(value = if_else(well_id %in% pos_ctrls & library_id == "KIN1", "pos_ctrl", value)) %>%
   mutate(type = if_else(well_id %in% pos_ctrls & library_id == "KIN1", "pos_ctrl", type)) %>%
  mutate(product_name = if_else(well_id %in% pos_ctrls & library_id == "KIN1","PF-670462", product_name)) %>% 
  mutate(catalog_number = if_else(well_id %in% pos_ctrls & library_id == "KIN1", "S6734", catalog_number)) %>% 
  mutate(target_cc= if_else(well_id %in% pos_ctrls & library_id == "KIN1", "CK1", target_cc)) %>%
    mutate(target_selleck= if_else(well_id %in% pos_ctrls & library_id == "KIN1", "CK1", target_selleck)) 



```

# Define annotation columns to used during the following analysis

```{r,echo=F,warning=F,message=F}
anno_columns <- c("cell_line","catalog_number","product_name","value","target_cc","target_selleck",
                   "cell_line_id","screen_id","library_id","library_plate_id",
                   "concentration","well_id", "plate_id","type", "InChIKey","cc_status")
```


## QC
# Boxplots
```{r,echo=F,warning=F,message=F}
# Boxplots per plate with DMSO controls for cell_count
raw_data %>% filter(value == "DMSO_ctrl") %>% 
  unite(plate_id,library_id,screen_id,library_plate_id,concentration) %>% 
    ggplot(aes(x=plate_id,y=cell_count,color=cell_line)) + geom_boxplot() +
      theme(axis.text.x = element_text(angle = 45, hjust=1)) +
        ggtitle("Boxplots per plate with DMSO controls for cell_count")

# Boxplots per plate with DMSO controls for dot count
raw_data %>% filter(value == "DMSO_ctrl") %>% 
  unite(plate_id,library_id,screen_id,library_plate_id,concentration) %>% 
    ggplot(aes(x=plate_id,y=punct_count,color=cell_line)) + geom_boxplot() + 
        theme(axis.text.x = element_text(angle = 45, hjust=1)) +
          ggtitle("Boxplots per plate with DMSO controls for dot count")


# Boxplots per plate with all samples for cell_count
raw_data %>% filter(value == "sample") %>% 
  unite(plate_id,library_id,screen_id,library_plate_id,concentration) %>% 
    ggplot(aes(x=plate_id,y=cell_count,color=cell_line)) + geom_boxplot() + 
        theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
          ggtitle("Boxplots per plate with samples for dot count")

# Boxplots per plate with all samples for dot count
raw_data %>% filter(value == "sample") %>% 
  unite(plate_id,library_id,screen_id,library_plate_id,concentration) %>% 
    ggplot(aes(x=plate_id,y=punct_count,color=cell_line)) + geom_boxplot() + 
        theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
          ggtitle("Boxplots per plate with samples for dot count")

# Dot plot per plate with positive controls

raw_data %>% 
  filter(value %in% c("DMSO_ctrl","pos_ctrl")) %>% 
  unite(plate_id,library_id,screen_id,library_plate_id,concentration) %>% 
    ggplot(aes(x=plate_id,y=punct_count,color=value)) + 
          geom_boxplot() + 
            facet_wrap(~cell_line) + 
              theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
ggtitle("Dot plot per plate with samples for dot count with DMSO boxplots overlayed")

```


# Plate plots
```{r,echo=F,warning=F,message=F}
raw_data %>% filter(cell_line == "wildtype") %>% 
  unite(plate_id,library_id,library_plate_id,concentration) %>% 
        separate(well_id,c("row","column"),1) %>% 
  mutate(column = factor(as.numeric(column))) %>% 
      mutate(row = factor(row,levels = rev(LETTERS[1:16]), labels = as.factor(rev(LETTERS[1:16])))) %>% 
        ggplot(aes(x=column,y=row)) + geom_tile(aes(fill=cell_count), colour = "white") + 
          scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu"))) + facet_wrap(~plate_id+screen_id) + 
            ggtitle("plate plots wildtype for cell count")

raw_data %>% filter(cell_line == "Dsh_eos") %>% 
  unite(plate_id,library_id,library_plate_id,concentration) %>% 
        separate(well_id,c("row","column"),1) %>% 
  mutate(column = factor(as.numeric(column))) %>% 
      mutate(row = factor(row,levels = rev(LETTERS[1:16]), labels = as.factor(rev(LETTERS[1:16])))) %>% 
        ggplot(aes(x=column,y=row)) + geom_tile(aes(fill=cell_count), colour = "white") + 
          scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu"))) + facet_wrap(~plate_id+screen_id) + 
            ggtitle("plate plots Dsh_eos for cell count")

raw_data %>% filter(cell_line == "Dsh_eos") %>% 
  unite(plate_id,library_id,library_plate_id,concentration) %>% 
        separate(well_id,c("row","column"),1) %>% 
  mutate(column = factor(as.numeric(column))) %>% 
      mutate(row = factor(row,levels = rev(LETTERS[1:16]), labels = as.factor(rev(LETTERS[1:16])))) %>% 
        ggplot(aes(x=column,y=row)) + geom_tile(aes(fill=punct_count), colour = "white") + 
          scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu"))) + facet_wrap(~plate_id) + 
            ggtitle("plate plots Dsh_eos for dots")


# flag well with high count
raw_data %>% filter(cell_line == "Dsh_eos") %>% 
  unite(plate_id,library_id,library_plate_id,concentration) %>% 
        separate(well_id,c("row","column"),1) %>% 
  mutate(column = factor(as.numeric(column))) %>% 
  mutate(punct_count = if_else(punct_count>100,NaN,punct_count)) %>% 
      mutate(row = factor(row,levels = rev(LETTERS[1:16]), labels = as.factor(rev(LETTERS[1:16])))) %>% 
        ggplot(aes(x=column,y=row)) + geom_tile(aes(fill=punct_count), colour = "white") + 
          scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu"))) + facet_wrap(~plate_id+screen_id) + 
            ggtitle("plate plots Dsh_eos for dots")

```


# Only keep the Dsh-eos line
```{r,echo=F,warning=F,message=F}

raw_data_eos <- raw_data %>% filter(cell_line == "Dsh_eos")

```


# Flag data points with high cell count (which are probably contaminated)

Could be done more in a elegant way by setting a quantile cutoff

```{r,echo=F,warning=F,message=F}
raw_data_eos %>% filter(cell_line == "Dsh_eos" & value == "DMSO_ctrl") %>%
  ggplot(aes(x=cell_count)) + geom_histogram()

#raw_data_eos <- raw_data_eos %>% 
    #  mutate(cell_count = if_else(cell_count > 1400,NaN, cell_count)) %>% 
    #  na.omit() %>%
    #  filter(target!="DMSO_ctrl_phenolRed")
```

# Glog transform that data 

```{r,echo=F,warning=F,message=F}
for(i in colnames(raw_data_eos[,!colnames(raw_data_eos) %in% anno_columns])) {
    q = quantile(raw_data_eos[,i],probs=0.05,na.rm=T)
    if(q == 0){
        q = 0.05 * max(raw_data_eos[,i],na.rm = T)
    }
    raw_data_eos[,i] <- glog(raw_data_eos[,i], q)
}

```

# Bring the data in long format

```{r,echo=F,warning=F,message=F}
raw_data_long <- raw_data_eos %>% 
  melt(id.vars = anno_columns,
       variable.name = "feature", value.name = "measurement") %>%
        as_tibble()
```



# Normalization 
We first normalize per plate and then the dots relative to the cell count per well. 
We also keep the un-normalized puncta counts.

```{r,echo=F,warning=F,message=F}

norm_data <- raw_data_long %>% 
  group_by(feature,library_id,library_plate_id,screen_id,concentration,cell_line) %>%
  mutate(measurement = measurement - median(measurement,na.rm = T)) %>%
   ungroup() 

norm_data_punct_count <- norm_data %>% filter(feature == "punct_count")

norm_data <- norm_data %>% 
                  group_by(library_id,library_plate_id,screen_id,concentration,cell_line,well_id) %>%
                  mutate(measurement = if_else(feature == "punct_count",
                                             measurement - measurement[feature == "cell_count"],
                                             measurement)) %>% 
                  ungroup() %>% 
                  mutate(feature = if_else(feature ==  "punct_count",
                                                          "punct_count_norm",
                                                         as.character(feature))) %>% 
                    bind_rows(norm_data_punct_count)

norm_data <- norm_data %>%
  group_by(feature) %>%
  mutate(measurement=measurement/mad(measurement[value=="DMSO_ctrl"],na.rm=T))

norm_data_spread <- norm_data %>%
  filter(feature %in% c(
    "punct_count","punct_count_norm",
    "cell_count","mean_cell.0.m.eccentricity",
    "mean_cell.0.s.area","mean_nuc.0.s.area")
    ) %>%
  spread(feature,measurement) 



replicate_spread <- norm_data_punct_count %>% spread(screen_id,measurement) %>% 
                      mutate(type = if_else(product_name=="DMSO_ctrl","neg",if_else(product_name=="PF-670462","pos","sample"))) %>%
  filter(concentration=="h")

p1 <- replicate_spread %>% 
  ggplot(aes(x=S02,y=S03)) + 
    geom_point(data = replicate_spread %>% filter(type=="sample"),col="darkgrey") +
    geom_point(data = replicate_spread %>% filter(type=="neg"),col="#f653a6") +
    geom_point(data = replicate_spread %>% filter(type=="pos"),col="blue") +
    scale_color_manual(values = c("#f653a6","blue","darkgrey")) +
    geom_smooth(method = "lm",show.legend = T,data = replicate_spread,col="grey") +
    ggtitle(paste0("PCC = ",round(cor(replicate_spread$S02,replicate_spread$S03,use = "p",method = "p"),digits = 2)))

p2 <- replicate_spread %>% 
  ggplot(aes(x=S01,y=S03)) + 
    geom_point(data = replicate_spread %>% filter(type=="sample"),col="darkgrey") +
    geom_point(data = replicate_spread %>% filter(type=="neg"),col="#f653a6") +
    geom_point(data = replicate_spread %>% filter(type=="pos"),col="blue") +
    scale_color_manual(values = c("#f653a6","blue","darkgrey")) +
    geom_smooth(method = "lm",show.legend = T,data = replicate_spread,col="grey") +
    ggtitle(paste0("PCC = ",round(cor(replicate_spread$S01,replicate_spread$S03,use = "p",method = "p"),digits = 2)))

p3 <- replicate_spread %>% 
  ggplot(aes(x=S01,y=S02)) + 
    geom_point(data = replicate_spread %>% filter(type=="sample"),col="darkgrey") +
    geom_point(data = replicate_spread %>% filter(type=="neg"),col="#f653a6") +
    geom_point(data = replicate_spread %>% filter(type=="pos"),col="blue") +
    scale_color_manual(values = c("#f653a6","blue","darkgrey")) +
    geom_smooth(method = "lm",show.legend = T,data = replicate_spread,col="grey") +
    ggtitle(paste0("PCC = ",round(cor(replicate_spread$S02,replicate_spread$S01,use = "p",method = "p"),digits = 2)))

ggsave(filename = "replicate_correlation_main_high.pdf",plot = p1,height = 8,width = 8)

ggsave(filename = "replicate_correlation_supplement_high.pdf",plot = p1 + p2 +p3,height = 8,width = 24)


replicate_spread <- norm_data_punct_count %>% spread(screen_id,measurement) %>% 
                      mutate(type = if_else(product_name=="DMSO_ctrl","neg",if_else(product_name=="PF-670462","pos","sample"))) %>%
  filter(concentration=="l")

p1 <- replicate_spread %>% 
  ggplot(aes(x=S02,y=S03)) + 
    geom_point(data = replicate_spread %>% filter(type=="sample"),col="darkgrey") +
    geom_point(data = replicate_spread %>% filter(type=="neg"),col="#f653a6") +
    geom_point(data = replicate_spread %>% filter(type=="pos"),col="blue") +
    scale_color_manual(values = c("#f653a6","blue","darkgrey")) +
    geom_smooth(method = "lm",show.legend = T,data = replicate_spread,col="grey") +
    ggtitle(paste0("PCC = ",round(cor(replicate_spread$S02,replicate_spread$S03,use = "p",method = "p"),digits = 2)))

p2 <- replicate_spread %>% 
  ggplot(aes(x=S01,y=S03)) + 
    geom_point(data = replicate_spread %>% filter(type=="sample"),col="darkgrey") +
    geom_point(data = replicate_spread %>% filter(type=="neg"),col="#f653a6") +
    geom_point(data = replicate_spread %>% filter(type=="pos"),col="blue") +
    scale_color_manual(values = c("#f653a6","blue","darkgrey")) +
    geom_smooth(method = "lm",show.legend = T,data = replicate_spread,col="grey") +
    ggtitle(paste0("PCC = ",round(cor(replicate_spread$S01,replicate_spread$S03,use = "p",method = "p"),digits = 2)))

p3 <- replicate_spread %>% 
  ggplot(aes(x=S01,y=S02)) + 
    geom_point(data = replicate_spread %>% filter(type=="sample"),col="darkgrey") +
    geom_point(data = replicate_spread %>% filter(type=="neg"),col="#f653a6") +
    geom_point(data = replicate_spread %>% filter(type=="pos"),col="blue") +
    scale_color_manual(values = c("#f653a6","blue","darkgrey")) +
    geom_smooth(method = "lm",show.legend = T,data = replicate_spread,col="grey") +
    ggtitle(paste0("PCC = ",round(cor(replicate_spread$S02,replicate_spread$S01,use = "p",method = "p"),digits = 2)))

ggsave(filename = "replicate_correlation_main_low.pdf",plot = p1,height = 8,width = 8)

ggsave(filename = "replicate_correlation_supplement_low.pdf",plot = p1 + p2 +p3,height = 8,width = 24)

p1 <- norm_data_spread %>% 
  filter(product_name %in% c("Afatinib (BIBW2992)","AT9283","Silmitasertib (CX-4945)","BMS-754807","PF-670462","SC1","DMSO_ctrl")) %>%
  ggplot(aes(x=reorder(product_name,punct_count_norm,median),y=punct_count_norm)) +
    geom_boxplot_jitter(fill="darkgrey") +
    facet_wrap(~concentration)+ 
  theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust=0.1))+
    xlab("compound") +
    ylab("puncta count [z]")

ggsave(filename = "treatment_boxplots.pdf",plot = p1,height = 8,width = 24)

for(i in c("Afatinib (BIBW2992)","AT9283","Silmitasertib (CX-4945)","BMS-754807","PF-670462","SC1")){
  p1 <- norm_data_spread %>% 
  filter(product_name %in% c(i,"DMSO_ctrl")) %>%
  ggplot(aes(x=factor(product_name,levels=c(i,"DMSO_ctrl")),y=punct_count_norm)) +
    geom_boxplot_jitter(fill="darkgrey") +
    facet_wrap(~concentration)+ 
  #theme(axis.text.x = element_text(angle = -45, vjust = 0.5, hjust=0.1))+
    xlab("compound") +
    ylab("puncta count [z]")+
    ggtitle(i,subtitle = "p-value by Wilcoxon rank sum test")+
    stat_compare_means(comparisons = list(c("DMSO_ctrl",i)),label=NA)+
    ylim(-13,9)
  
  ggsave(filename = paste0(i,"_treatment_boxplots.pdf"),plot = p1,height = 6,width = 4)

  
}
    

a <- norm_data_spread %>% 
  group_by(catalog_number) %>%
  summarise_if(is.numeric,mean) %>%
  ggplot(aes(x=cell_count,y=punct_count_norm)) +
    geom_point_rast(raster.dpi = 300,raster.width = 4,raster.height = 4) +
    geom_smooth(method = "lm") +
    xlab("cell count [z-score]") +
    ylab("puncta count [z-score]")

b <- norm_data_spread %>% 
  group_by(catalog_number) %>%
  summarise_if(is.numeric,mean) %>%
  ggplot(aes(x=mean_cell.0.s.area,y=punct_count_norm)) +
    geom_point_rast(raster.dpi = 300,raster.width = 4,raster.height = 4) +
    geom_smooth(method = "lm") +
    xlab("cell area [z-score]") +
    ylab("puncta count [z-score]")

c <- norm_data_spread %>% 
  group_by(catalog_number) %>%
  summarise_if(is.numeric,mean) %>%
  ggplot(aes(x=mean_cell.0.m.eccentricity,y=punct_count_norm)) +
    geom_point_rast(raster.dpi = 300,raster.width = 4,raster.height = 4) +
    geom_smooth(method = "lm") +
    xlab("cell eccentricity [z-score]") +
    ylab("puncta count [z-score]")

d <- norm_data_spread %>% 
  group_by(catalog_number) %>%
  summarise_if(is.numeric,mean) %>%
  ggplot(aes(x=mean_nuc.0.s.area,y=punct_count_norm)) +
    geom_point_rast(raster.dpi = 300,raster.width = 4,raster.height = 4) +
    geom_smooth(method = "lm") +
    xlab("nuclear size [z-score]") +
    ylab("puncta count [z-score]")

require(patchwork)
p <- a+b+c+d
ggsave("results/morphology_vs_condensates_per_cell.pdf",p,width = 8,height = 8)


```

# Hit calling

```{r,echo=F,warning=F,message=F}
# we first have a look at the distribution of the dot counts 
norm_data %>% 
filter(feature == "punct_count_norm"  & 
         value != "DMSO_ctrl_phenolRed") %>% 
    group_by(product_name,concentration,target_cc) %>%
  summarise(measurement=mean(measurement)) %>%
  ungroup() %>%
ggplot(aes(x=measurement)) + geom_histogram() + 
  xlab("Punct count per cell (well/screen average) [z-score]") +
    ggtitle("Histogram for Dsh-eos")

# Unfortunately stat_qq does not support labels so we re-code the QQ plot
sample <- norm_data %>% 
  filter(feature == "punct_count_norm" & cell_line == "Dsh_eos" & value != "DMSO_ctrl_phenolRed") %>% 
  group_by(product_name,concentration,target_cc) %>%
  summarise(measurement=mean(measurement)) %>%
  ungroup() %>%
  na.omit()  %>% 
  arrange(measurement) %>% 
  mutate(rank = seq(1:nrow(.))) 

labels <- sample %>% filter(rank<=10 | rank>= nrow(.)-10) %>% 
            unite(label,product_name,concentration) %>% 
              select(label,rank)

probabilities = (1:nrow(sample))/(nrow(sample)+1)

theoretical <- qnorm(probabilities, mean(sample$measurement),
                     sd(sample$measurement))

sample_theoretical <- sample %>% arrange(measurement) %>% 
                        mutate(theoretical = sort(theoretical))

labels <- labels %>% left_join(sample_theoretical,by=c("rank"))

ref_line <- cbind.data.frame(
  standard=qnorm(c(0.25,0.75)),
  empirical=quantile(sample_theoretical$theoretical, c(0.25,0.75)))

ref_inter <- lm(standard~empirical,data = ref_line)$coefficients[1]
ref_slope <- lm(standard~empirical,data = ref_line)$coefficients[2]

sample_theoretical %>% 
  ggplot(aes(x=theoretical,y=measurement)) + 
    geom_point() + geom_text_repel(data=labels,aes(x=theoretical,y=measurement,label=label)) + 
            theme(aspect.ratio=1) +  
  geom_abline(slope = ref_slope,intercept = ref_inter) + 
  ylab("sample quantiles") + xlab("theoretical quantiles") + 
          ggtitle("Q-Q plot for the punct count per cell (well/screen average) [z-scores]")


# with targets instead of compound names
labels <- sample %>% filter(rank<=10 | rank>= nrow(.)-10) %>% 
            mutate(label = target_cc) %>% 
              select(label,rank)

sample_theoretical <- sample %>% arrange(measurement) %>% 
  mutate(theoretical = sort(theoretical))


labels <- labels %>% left_join(sample_theoretical,by=c("rank"))

p <-sample_theoretical %>% 
 ggplot(aes(x=theoretical,y=measurement)) + 
      geom_point_rast(raster.width = 8,raster.height = 8,raster.dpi = 400) +
      geom_text_repel(data=labels,aes(x=theoretical,y=measurement,label=label)) +
    geom_abline(slope = ref_slope,intercept = ref_inter) + 
    theme(aspect.ratio=1) + 
    ylab("sample quantiles") + 
    xlab("theoretical quantiles") + 
    ggtitle("Q-Q plot for the punct count per cell (well average) [z-scores]")+
    facet_wrap(~concentration)

print(p)
 
ggsave("results/QQplot.pdf",p,width = 16,height = 8)
```

# Calculate statistical effect size between positive controls and DMSO contorls 

```{r,echo=F,warning=F,message=F}
norm_data %>% filter(value %in% c("pos_ctrl","DMSO_ctrl")) %>% 
  filter(feature == "punct_count_norm") %>% 
  ggplot(aes(x=product_name,y=measurement)) + 
    facet_wrap(~concentration) +
    stat_compare_means(comparisons = list(c("DMSO_ctrl","PF-670462")),label=NA) +
    geom_boxplot() + xlab("sampe type") + 
    ylab("punct count per cell (well average) [z-score]") +
    ggtitle("Statistical effect size between contorls for Dsh-eos line (Wilcoxon rank sum test)")

ggsave("results/Control_difference.pdf",width = 8,height = 8)
```

#Statistical analysis

```{r}

essentials <- 
  norm_data %>% 
  filter(feature=="cell_count") %>% 
  group_by(product_name) %>% 
  summarise(essentiality=mean(measurement))

nuclear_size <- norm_data %>% 
  filter(feature=="mean_nuc.0.s.area") %>% 
  group_by(product_name) %>% 
  summarise(nuc_size=mean(measurement))


normed_data <- norm_data %>% 
  filter(feature == "punct_count_norm" & cell_line == "Dsh_eos" & value != "DMSO_ctrl_phenolRed"& value != "DMSO_ctrl") %>% 
  group_by(product_name,catalog_number,concentration,target_cc,screen_id) %>%
  summarise(measurement=mean(measurement)) %>%
  ungroup() %>%
  na.omit() %>%
  spread(screen_id,measurement) %>%
  drop_na()

normed_data$pval <- normed_data %>% select_if(is.numeric) %>% limma::lmFit() %>% limma::eBayes() %>% .$p.value %>% as.numeric()

normed_data <- normed_data %>% 
                  mutate(fdr=p.adjust(pval,method = "BH")) %>% 
                  mutate(mean_z=(S01+S02+S03)/3) %>%
                  left_join(essentials) %>%
                  left_join(nuclear_size)

normed_data %>% 
  filter(fdr<0.2) %>%
  rename(Screen_1_puncta_z=S01,Screen_2_puncta_z=S02,Screen_3_puncta_z=S03,Avg_puncta_z=mean_z,normed_cell_count_z=essentiality,normed_nucleus_area=nuc_size) %>%
  write_delim(file = "results/hit_list_fdr_0.2.csv",delim = ",")

normed_data %>% 
  rename(Screen_1_puncta_z=S01,Screen_2_puncta_z=S02,Screen_3_puncta_z=S03,Avg_puncta_z=mean_z,normed_cell_count_z=essentiality,normed_nucleus_area=nuc_size) %>%
  write_delim(file = "results/hit_list_unfiltered.csv",delim = ",")


p1<-normed_data %>%
  mutate(concentration=if_else(concentration=="l","low concentration","high concentration")) %>%
  ggplot(aes(x=mean_z,y = -log10(pval),label=product_name)) +
    #geom_point_rast(raster.dpi = 400) +
    geom_point() +
    geom_label_repel(data = normed_data %>%
  mutate(concentration=if_else(concentration=="l","low concentration","high concentration")) %>% filter(fdr<0.1) %>% slice_max(order_by=abs(mean_z),n=20))+
  facet_wrap(~concentration) +
    ylab("-log10 p-value") +
    theme(aspect.ratio=1) + 
    xlab("puncta count per cell [z-score]")

print(p1)    

p2<-normed_data %>%
  mutate(concentration=if_else(concentration=="l","low concentration","high concentration")) %>%
  ggplot(aes(x=mean_z,y = -log10(pval),label=target_cc)) +
    #geom_point_rast(raster.dpi = 400) +
    geom_point() +
    geom_label_repel(data = normed_data %>%
  mutate(concentration=if_else(concentration=="l","low concentration","high concentration")) %>% filter(fdr<0.1) %>% slice_max(order_by=abs(mean_z),n=20))+
  facet_wrap(~concentration) +
    ylab("-log10 p-value") +
    theme(aspect.ratio=1) + 
    xlab("puncta count per cell [z-score]")

print(p2)    

p3<-normed_data %>%
  mutate(concentration=if_else(concentration=="l","low concentration","high concentration")) %>%
  ggplot(aes(x=mean_z,y = -log10(pval),label=target_cc)) +
    #geom_point_rast(raster.dpi = 400) +
    geom_point() +
    geom_label_repel(data = normed_data %>%  filter(fdr<0.1) %>% slice_max(order_by=abs(mean_z),n=20))+
    ylab("-log10 p-value") +
    theme(aspect.ratio=1) + 
    xlab("puncta count per cell [z-score]")

print(p3)    


ggsave("results/Volcanoplot_product_names.pdf",p1,width = 16,height = 8)
ggsave("results/Volcanoplot_targets.pdf",p2,width = 16,height = 8)
ggsave("results/Volcanoplot_combined.pdf",p3,width = 16,height = 8)


# Volcano plot to highlight CFI-400945
normed_data %>%
  mutate(concentration=if_else(concentration=="l","low concentration","high concentration")) %>%
  ggplot(aes(x=mean_z,y = -log10(pval),label=product_name)) +
    geom_point() +
   geom_label_repel(
     data = filter(normed_data,product_name == "CFI-400945") %>% 
       mutate(concentration=if_else(concentration=="l","low concentration","high concentration")),
                  nudge_y       = 1,
                  size          = 4
   )+
  facet_wrap(~concentration) +
    xlab("puncta count per cell [z-score]") +
    ylab("-log10 p-value")


normed_data %>%
  mutate(concentration=if_else(concentration=="l","low concentration","high concentration")) %>%
  ggplot(aes(x=mean_z,y = -log10(pval),label=product_name)) +
    geom_point() +
   geom_label_repel(
     data = filter(normed_data,product_name %in% unique(norm_data$product_name[grepl("PLK",norm_data$target_cc)])) %>% 
       mutate(concentration=if_else(concentration=="l","low concentration","high concentration")),
                  nudge_y       = 1,
                  size          = 4
   )+
  facet_wrap(~concentration) +
    xlab("puncta count per cell [z-score]") +
    ylab("-log10 p-value")

```


# Session info
```{r}
sessionInfo()
```