---
title: "Script to document the image analysis for HC1098"
author:  
- "Christian Scheeder"
- "Division Signalling and Functional Genomics"
- "German Cancer Research Center (DKFZ) Heidelberg"
date: \today
output:
    BiocStyle::pdf_document
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
editor_options: 
    chunk_output_type: console
---

# Preamble
This script documents the image analysis for the project HC1098 where 
HEK 293 cells with endogenously tagged Dishevelled (Dsh) were screened against 
the selleck kinase library. The image analysis parameters defined here 
were transferred to a pipeline. 


# Session Setup    
Load the required packages
```{r, echo=F, results='hide', warning=F, message=F}
library(stringr)
library(dplyr)
library(dbplyr)
library(here)
library(EBImage)
library(parallel)
library(FNN)
library(RPostgreSQL)
```



# Load images of control wells
Note: The image analysis is set based on negative (DMSO only) and positive 
controls (5 µM PF-670462, target = CK1 $\epsilon$ and $\delta$).


```{r,fig.align="center"}
# read the DMSO images
DMSO_DAPI <- readImage(here("images/DMSO","H_12_f1_c2_xDAPI_mDAPI_z1_t1.tif")) 
DMSO_cy5 <- readImage(here("images/DMSO","H_12_f1_c3_xCy5_mCy5_z1_t1.tif"))
DMSO_FITC <- readImage(here("images/DMSO","H_12_f1_c1_xFITC_mFITC_z1_t1.tif"))

# let's have a brief look at the DMSO images
display(combine(
  normalize(DMSO_DAPI)*5,
  normalize(DMSO_cy5)*5,
  normalize(DMSO_FITC)*5),all=T,method = 'raster')
text(500,200, label="DAPI", col="red", cex=2)
text(2500,200, label="Cy5", col="red", cex=2)
text(500,2300, label="FITC", col="red", cex=2)


# read the PF-670462 images
pos_DAPI <- readImage(here("images/PF-670462","G_12_f1_c2_xDAPI_mDAPI_z1_t1.tif")) 
pos_cy5 <- readImage(here("images/PF-670462","G_12_f1_c3_xCy5_mCy5_z1_t1.tif"))
pos_FITC <- readImage(here("images/PF-670462","G_12_f1_c1_xFITC_mFITC_z1_t1.tif"))

# let's have a brief look at the PF-670462 images
display(combine(
  normalize(pos_DAPI)*5,
  normalize(pos_cy5)*5,
  normalize(pos_FITC)*5),all=T,method = 'raster')
text(500,200, label="DAPI", col="red", cex=2)
text(2500,200, label="Cy5", col="red", cex=2)
text(500,2300, label="FITC", col="red", cex=2)
```


# Segment the cells based on the DNA and Actin staining
```{r,fig.align="center"}
## blur images with gaussian filter 
DMSO_DAPI_smooth = gblur(DMSO_DAPI, radius = 51, sigma = 1)
DMSO_cy5_smooth = gblur(DMSO_cy5, radius  = 51, sigma = 4)


pos_DAPI_smooth = gblur(pos_DAPI, radius = 51, sigma = 1)
pos_cy5_smooth = gblur(pos_cy5, radius  = 51, sigma = 4)



# segment nuclei using adaptive thresholding 
#  the window sized is chosen based on the proximal size of the nuclei 
#      the first threshold is a bit smaller to sparate the nuclei
#      the second is larger to fill gaps 
# both masks are then combined using a propagation 

DMSO_nucleusTresh = thresh(DMSO_DAPI_smooth, 
                          w = 20, h = 20, 
                          offset = 0.004)

DMSO_nucleusTresh = fillHull(opening(DMSO_nucleusTresh,
                           kern=makeBrush(9, shape="disc")))

DMSO_nucleusFill = fillHull(thresh(DMSO_DAPI_smooth, 
                          w = 30, h = 30, 
                          offset = 0.001))

DMSO_nucleusRegions = propagate(DMSO_DAPI_smooth,
                           seed=bwlabel(DMSO_nucleusTresh),
                           mask=DMSO_nucleusFill)

# print the nuceli mask 
display(colorLabels(DMSO_nucleusRegions),method = 'raster')

# same for the positive ctrls
pos_nucleusTresh = thresh(pos_DAPI_smooth, 
                          w = 20, h = 20, 
                          offset = 0.004)

pos_nucleusTresh = fillHull(opening(pos_nucleusTresh,
                           kern=makeBrush(9, shape="disc")))

pos_nucleusFill = fillHull(thresh(pos_DAPI_smooth, 
                          w = 30, h = 30, 
                          offset = 0.001))

pos_nucleusRegions = propagate(pos_DAPI_smooth,
                           seed=bwlabel(pos_nucleusTresh),
                           mask=pos_nucleusFill)

## segment cell bodies by adaptive and global thersholding 
##      the actin channel is used by combinig an adaptive and global threshold
##      regions with mising cytoplasm signal are filled with 
##          the nucleus binary mask and a fillHull command

DMSO_cytoplasmThresh = thresh(DMSO_cy5_smooth,
                         w = 28,
                         h = 28,
                         offset =  0.001)

DMSO_cytoplasmOpening = opening(DMSO_cytoplasmThresh,
                           kern=makeBrush(9,shape="disc"))


DMSO_cytoplasmOpening2 = opening(DMSO_cy5_smooth > 0.03)


DMSO_nucleusRegions2 <- DMSO_nucleusRegions
DMSO_nucleusRegions2[DMSO_nucleusRegions != 0] <- 1


DMSO_cytoplasmCombined = DMSO_cytoplasmOpening | DMSO_cytoplasmOpening2 |DMSO_nucleusRegions2 


DMSO_cytoplasmRegions = propagate(x = DMSO_cy5_smooth,
                             seeds = DMSO_nucleusRegions,
                             lambda=1e-04,
                             mask=DMSO_cytoplasmCombined)

DMSO_cytoplasmRegions = fillHull(DMSO_cytoplasmRegions)

# print the cell body mask
display(colorLabels(DMSO_cytoplasmRegions),method = 'raster')

# generate an RBG oberlay and add the segmentation outlines 
DMSO_ImgColor = rgbImage(5*normalize(DMSO_cy5),
                    5*normalize(DMSO_FITC),
                    5*normalize(DMSO_DAPI))
            
DMSO_ImgOut = paintObjects(DMSO_cytoplasmRegions,
                      paintObjects(DMSO_nucleusRegions,
                                   DMSO_ImgColor,
                                   col='blue'),
                      col='white')

# Print the RGB overlay with segmentation borders
display(DMSO_ImgOut,method='raster')

# repeat the cell body steps for the positive control image 
pos_cytoplasmThresh = thresh(pos_cy5_smooth,
                         w = 28,
                         h = 28,
                         offset =  0.001)

pos_cytoplasmOpening = opening(pos_cytoplasmThresh,
                           kern=makeBrush(9,shape="disc"))


pos_cytoplasmOpening2 = opening(pos_cy5_smooth > 0.025)


pos_nucleusRegions2 <- pos_nucleusRegions
pos_nucleusRegions2[pos_nucleusRegions != 0] <- 1


pos_cytoplasmCombined = pos_cytoplasmOpening | pos_cytoplasmOpening2 | pos_nucleusRegions2 


pos_cytoplasmRegions = propagate(x = pos_cy5_smooth,
                             seeds = pos_nucleusRegions,
                             lambda=1e-04,
                             mask=pos_cytoplasmCombined)

pos_cytoplasmRegions = fillHull(pos_cytoplasmRegions)

pos_ImgColor = rgbImage(5*normalize(pos_cy5),
                    5*normalize(pos_FITC),
                    5*normalize(pos_DAPI))
            
pos_ImgOut = paintObjects(pos_cytoplasmRegions,
                      paintObjects(pos_nucleusRegions,
                                   pos_ImgColor,
                                   col='blue'),
                      col='white')


display(combine(DMSO_ImgOut,pos_ImgOut),
        method = 'raster',
        all=T)
text(500,200, label="DMSO", col="white", cex=2)
text(2900,200, label="PF-670462", col="white", cex=2)
```


# Segment objects in the FITC channel 
Two appraoches are compared: either a global threshold or an adative threshold 
combined with an opening. The global threshold is more 
conservative and thus preferred. 
```{r,fig.align="center"}
# We start with a DMSO image

# smoothen the images 
DMSO_FITC_smooth = gblur(DMSO_FITC, radius  = 51, sigma = 4)
pos_FITC_smooth = gblur(pos_FITC, radius  = 51, sigma = 4)


# global threshold (a simply cutoff at a chosen intensity value)
DMSO_puncTresh_global <- DMSO_FITC_smooth>0.025

# adaptive threshold 
#   (the thershold adapts to the mean intensity at each region of the image)
DMSO_puncTresh_adapt = thresh(DMSO_FITC_smooth, 
                          w = 20, h = 20, 
                          offset = 0.006)

# opening to remove tiny objects after the adaptive thresholding
DMSO_puncTresh_adapt = fillHull(opening(DMSO_puncTresh_adapt,
                           kern=makeBrush(7, shape="disc")))


# compare the two approaches 
display(combine(colorLabels(bwlabel(DMSO_puncTresh_global)),
        colorLabels(bwlabel(DMSO_puncTresh_adapt))),
        method = 'raster',
        all=T)

text(500,200, label="global", col="white", cex=2)
text(2900,200, label="adaptive", col="white", cex=2)



# global thresholding for positive contorl 
pos_puncTresh_global <- pos_FITC_smooth>0.025


# compare DMSO and positive ctrl
display(combine(colorLabels(bwlabel(DMSO_puncTresh_global)),
        colorLabels(bwlabel(pos_puncTresh_global))),
        method = 'raster',
        all=T)

text(500,200, label="DMSO", col="white", cex=2)
text(2900,200, label="PF-670462", col="white", cex=2)


# only the number of objects is kept as a feature
DMSO_n_punc <- length(unique(bwlabel(DMSO_puncTresh_global)[bwlabel(DMSO_puncTresh_global)!=0]))
pos_n_punc <- length(unique(bwlabel(pos_puncTresh_global)[bwlabel(pos_puncTresh_global)!=0]))

# print results
print(paste0("Number of objects identified in the DMSO image: ",DMSO_n_punc))
print(paste0("Number of objects identified in the PF-670462 image: ",pos_n_punc))
```


# Final notes
We will calculate the cell morphology features based on the 
DNA and Actin staining.  For the FITC channel the intensity featuress 
inside cells will be calculated (e.g. mean intensity). 
In addition we will save the number of objects identified in the the FITC channel. 


# Session info
```{r}
sessionInfo()
```



